EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-06-02 01_intro.tex
v1.1 arXiv ready AA, MC and GB

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems, capturing the coupling of a small number of localized degrees of freedom to an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have led to the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} have progressively become powerful tools for exploring a wide range of quantum many-body phenomena \cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their importance has expanded beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the  impurity  can be defined more generally, so as to include orbital and spin local degrees of freedom as well as a finite number of lattice sites. 
This setup captures many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.
Most notably, quantum impurity problems lie at the core of many quantum embedding methods, like density-matrix embedding theory (DMET) \cite{Knizia2012PRL}, Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Variational Cluster Approximations \cite{Potthoff2003TEPJBCMACS,Senechal2008,Potthoff2011ACP,Nuss2011,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.
Within these approaches the complex physics of the lattice systems gets reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds and unconventional superconductors.

Accurately and efficiently solving a quantum impurity problem and computing the Green's functions, which are central in the DMFT framework, nonetheless remains a critical challenge, particularly for multi-orbital systems, systems
at low temperatures or endowed with lower symmetry.
%
To this end, over the years 
many advanced numerical techniques have been developed, incorporating cutting-edge strategies to reduce computational loads and leverage state-of-the-art algorithms \cite{Bauer2011JOSMTAE,Parcollet2015CPC}. 
These range from continuous-time quantum Monte Carlo 
methods \cite{Gull2011RMP,Rubtsov2005PRB,Haule2007PRB,Seth2016CPC,Wallerberger2019CPC},
to numerical renormalization group approaches \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Debertolis2021PRB} 
and density-matrix renormalization group \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Nunez-Fernandez2025A} or Configuration Interaction \cite{Zgid2012PRB,Lu2014PRB,Go2017PRB,Bi2019CPC,Mejuto-Zaera2019PRB} solvers. 
In this context, the Exact Diagonalization (ED) approach plays a significant 
and distinctive role \cite{Caffarel1994PRL,Dolfen2006,Perroni2007PRB,Capone2007PRB,Weber2012PRB,Lu2017TEPJST,Amaricci2022CPC}.

More generally, several open-source ED software packages have recently been made available  within the condensed matter community to address quantum many-body problems with an emphasis on efficiency and flexibility. Examples include \texttt{EDLib} \cite{Iskakov2017} and the newer 
\texttt{XDiag} \cite{Wietek2025}, which focus on generic fermionic systems; 
\texttt{Pomerol} \cite{Antipov2015} designed for fermion-boson models and interfaced with TRIQS \cite{Parcollet2015CPC}; \texttt{H$\Phi$} \cite{Kawamura2017CPC,Ido2024CPC} 
for fermions coupled to spins; and general-purpose spin-model solvers such as 
\texttt{QuSpin} \cite{Weinberg2017SP,Weinberg2019SP}, all contributing to the 
rich landscape of ED-based techniques.

%EDIPACK
In this work we present the improved and renewed \NAME{}: a flexible, high-performance numerical library designed to address these challenges, providing an exact diagonalization (ED) solver for generic Quantum Impurity Problems. 
The library builds on the foundations of
its predecessor \cite{Amaricci2022CPC}, featuring massively parallel Lanczos-based algorithms \cite{Lanczos1950JRNBSB,Lehoucq1998,Maschhoff1996} while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations \cite{Amaricci2022CPC,Capone2007PRB} for a wide set of impurity models, making it suitable for addressing a broad class of problems, from spin-orbit coupling in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

Additionally, \NAME offers access to the Fock space of
the impurity system and thus, for instance, the possibility to evaluate the impurity reduced density matrix (iRDM), enabling quantum information-inspired analyses providing insights into the entanglement properties, subsystem purity, and quantum correlations of interacting electron systems \cite{Walsh2021PNAS,Bellomia2024PRB}.
In the quantum embedding context, these features are increasingly recognized as essential to understand the emergent behavior of correlated matter, including quantum criticality and superconductivity.
% MC this last point is intrinsically DMFT (or embedding) and it has no connection with impurity physics. Right: I added an opening which should clarify the sentence.

This new \NAME version also includes support for systems with inequivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites.
% This capability makes it 
% possible to explore a wide range of inhomogeneous systems, including
% oxide interfaces or disordered systems, where
% translational symmetry is broken.
%MC? Repetition.. removed

%\paragraph{Interoperability}\\
With the growing importance of quantum impurity models in the study of correlated materials, standardization and interoperability across software suites has
become a critical necessity.
\NAME has been expressly designed following this guiding principle. 
While the library itself is implemented using modern Fortran constructs, it also provides an extensive set of C-bindings that enable integration with other programming languages,
be they compiled (C, C++), JIT-compiled (Julia) or interpreted (Python).  
Especially notable in this sense is the Python API, EDIpack2py, which beyond providing a 
high-level interface for defining and solving impurity problems, it also allows integration with established scientific libraries, such as NumPy and SciPy, facilitating data analysis and
rapid prototyping.

A key development feature in \NAME, unlocked by the introduction of the Python API, 
is the possibility to directly interface to both TRIQS \cite{Parcollet2015CPC} (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics \cite{Wallerberger2019CPC} software suites, which are two of the most important and widely used computational frameworks in the field.
\NAME slots in as an alternative solver to the native Quantum Monte Carlo 
implementations \cite{Gull2011RMP,Seth2016CPC}, extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the coding requirements for the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability enhances the collective reach and reliability of these platforms, 
improving reproducibility and cross-validation of numerical results, with the ultimate goal of achieving standardization for quantum impurity solvers along the guiding principles of initiatives such as FAIRmat (see \href{https://www.fairmat-nfdi.eu/fairmat/}{fairmat-nfdi.eu/fairmat}).

%\paragraph{closure}
In this paper, we present the mathematical framework, implementation
details, and key features of \NAME, along with several illustrative
applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging problems. 
The rest of this work is organized as follows. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration
and installation procedures. In \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface which is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) an experimental, yet operative, Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and  all the discussed extensions with a series of illustrative examples.

>>>>

\section{Introduction and Motivation}\label{SecIntro}
Quantum impurity models play a central role in the study of strongly correlated electron systems. They describe the coupling of a small number of localized degrees of freedom to an extended non-interacting environment \cite{Nozieres1980JP,Hewson1993}. 
Originally introduced to describe diluted magnetic impurities in metallic hosts, single-impurity models have played a central role in the field of quantum many-body systems through the decades, starting from the full understanding of the Kondo problem \cite{Anderson1961PR,Kondo1964POTP,Schrieffer1966PR} and subsequently proving crucial to the understanding of a wide range of quantum many-body phenomena\cite{Wilson1975RMP,Georges1996RMP,Kotliar2004PT,Kotliar2006RMP}. More recently, their scope has expanded beyond the realm of condensed matter, influencing the understanding of other areas such as quantum information \cite{Su2013MPLB,Walsh2019PRL,Walsh2020PQ,Walsh2021PNAS,Stocker2022,Bellomia2024PRB} or ultra-cold atomic systems \cite{Dao2007PRL,Amaricci2014PRA,Del-Re2018PRA,Walsh2019PRB,Tusi2022NP}.

At the heart of quantum impurity models is the interaction between a localized ``impurity''  and a surrounding bath of itinerant particles, usually represented as a conduction band of electrons \cite{Hewson1993}. In this context, the impurity  can be defined more generally, so as to include orbital and spin local degrees of freedom as well as a finite number of lattice sites, i.e. quantum clusters, immersed in a suitable environment. 
This setup lies at the core of several quantum embedding methods, including Density-Matrix Embedding Theory (DMET) \cite{Knizia2012PRL,Sun2020PRB} and further extensions \cite{Scott2021PRB,Nusspickel2022PRX}, the Gutzwiller Approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
Quantum Cluster methods \cite{Potthoff2003TEPJBCMACS,Potthoff2011ACP,Dionne2023SPC,Dionne2023SPCa}
or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}, capturing many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}.

%MC non in tutti questi metodi si introduce un modello di impurezza: 
%AA "sistemato" facendo una piccola generalizzazione del concetto di impurity problem, includendo quantum clusters (VCA, CPA, etc.) or embedded fragments (DMET). Accettabile? 
% This setup captures many essential aspects of strong correlation, including screening \cite{Roekeghem2014PRL,Roekeghem2014EL,Werner2016JOPCM,Tomczak2017TEPJST}, decoherence and entanglement in a computationally tractable way \cite{Walsh2021PNAS,Bellomia2024PRB}, lying at the core of several quantum embedding methods, such as Density-Matrix Embedding Theory (DMET) \cite{Knizia2012PRL,Sun2020PRB} and extensions\cite{Scott2021PRB,Nusspickel2022PRX}, the Gutzwiller approximation (GA) \cite{Lanata2015PRX,Mejuto-Zaera2023PRB},
% Quantum Cluster methods \cite{Potthoff2003TEPJBCMACS,Potthoff2011ACP,Dionne2023SPC,Dionne2023SPCa}
% or Dynamical Mean-Field Theory (DMFT) \cite{Georges1996RMP, Kotliar2004PT,Kotliar2006RMP}.

In several of these embedding schemes, most notably DMFT, the complex physics of the lattice systems gets reduced to an effective self-consistent quantum impurity problem.
By accounting for local quantum fluctuations, DMFT successfully describes the most important physical features of correlated materials, including Mott insulators, heavy fermion compounds and unconventional superconductors.  
%
Accurately and efficiently solving a quantum impurity problem and computing the Green's functions, which are central in the DMFT framework, nonetheless remains a critical challenge, particularly for multi-orbital systems, those 
at low temperatures or endowed with lower symmetry.
%
To this end, 
many advanced numerical techniques have been developed over the years, incorporating cutting-edge strategies to reduce computational loads and leverage state-of-the-art algorithms \cite{Bauer2011JOSMTAE,Parcollet2015CPC}. 
These range from continuous-time quantum Monte Carlo 
methods \cite{Gull2011RMP,Rubtsov2005PRB,Haule2007PRB,Seth2016CPC,Wallerberger2019CPC},
to numerical renormalization group approaches \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Debertolis2021PRB} 
and density-matrix renormalization group \cite{Zitko2009PRB,Bulla2001PRB,Bulla2008RMP,Nunez-Fernandez2025A} or Configuration Interaction \cite{Zgid2012PRB,Lu2014PRB,Go2017PRB,Bi2019CPC,Mejuto-Zaera2019PRB} solvers. 
In this context, the Exact Diagonalization (ED) approach plays a significant 
and distinctive role \cite{Caffarel1994PRL,Dolfen2006,Perroni2007PRB,Capone2007PRB,Weber2012PRB,Lu2017TEPJST,Amaricci2022CPC}.

Several open-source ED software packages have recently been made available within the condensed matter community to address quantum many-body problems with an emphasis on efficiency and flexibility. Examples include \texttt{EDLib} \cite{Iskakov2017} and the newer 
\texttt{XDiag} \cite{Wietek2025}, which focus on generic fermionic systems; 
\texttt{Pomerol} \cite{Antipov2015} designed for fermion-boson models and interfaced with TRIQS \cite{Parcollet2015CPC}; \texttt{H$\Phi$} \cite{Kawamura2017CPC,Ido2024CPC} 
for fermions coupled to spins; and general-purpose spin-model solvers such as 
\texttt{QuSpin} \cite{Weinberg2017SP,Weinberg2019SP}, all contributing to the 
rich landscape of ED-based techniques.

%EDIPACK
In this work we present the improved and renewed \NAME{}: a flexible, high-performance numerical library designed to address these challenges, providing an ED solver for generic quantum impurity problems. 
The library builds on the foundations of
its predecessor \cite{Amaricci2022CPC}, featuring massively parallel Lanczos-based algorithms \cite{Lanczos1950JRNBSB,Lehoucq1998,Maschhoff1996} while including several significant extensions and improvements. The library now supports, within a unified framework, both zero and finite temperature calculations \cite{Amaricci2022CPC,Capone2007PRB} for a wide set of impurity models, making it suitable for addressing a broad class of problems, from spin-orbit coupling in quantum materials to multi-orbital superconductivity. It also provides direct
access to numerically exact dynamical correlation functions across the
entire frequency complex plane, thus enabling calculations of spectral functions, local susceptibilities, and other response properties.

Additionally, \NAME offers access to the Fock space of
the impurity system and thus, for instance, the possibility to evaluate the impurity reduced density matrix, enabling quantum information-inspired analyses providing insights into the entanglement properties, subsystem purity, and quantum correlations of interacting electron systems \cite{Walsh2021PNAS,Bellomia2024PRB}.
In the quantum embedding context, these features are increasingly recognized as essential to understand the emergent behavior of correlated matter, including quantum criticality and superconductivity.

This new \NAME version also includes support for systems with inequivalent impurities. This is a critical feature in DMFT or, more generally, in quantum embedding methods, to study hetero-structures, super-lattices, and disordered systems, where the local environment varies significantly across different sites.

%\paragraph{Interoperability}\\
With the growing importance of quantum impurity models in the study of correlated materials, standardization and interoperability across software suites have become critical necessities.
\NAME has been expressly designed following this guiding principle. 
While the library itself is implemented using modern Fortran constructs, it also provides an extensive set of C-bindings that enable integration with other programming languages,
be they compiled (C, C++), JIT-compiled (Julia) or interpreted (Python).  
Especially notable in this sense is the Python API, EDIpack2py, which beyond providing a 
high-level interface for defining and solving impurity problems, it also allows integration with established scientific libraries, such as NumPy and SciPy, facilitating data analysis and
rapid prototyping.

A key development feature in \NAME, unlocked by the introduction of the Python API, 
is the possibility to directly interface to both TRIQS \cite{Parcollet2015CPC} (Toolbox for Research on Interacting Quantum Systems) and
W2Dynamics \cite{Wallerberger2019CPC} software suites, which are two of the most important and widely used computational frameworks in the field.
\NAME slots in as an alternative solver to the native Quantum Monte Carlo 
implementations \cite{Gull2011RMP,Seth2016CPC}, extending the parameter (and especially temperature) range reachable by 
the simulation while minimizing the coding requirements for the end user.
At the same time, this integration provides \NAME users with access to the
extensive data handling and analysis tools in the TRIQS ecosystem. 
Overall, interoperability enhances the collective reach and reliability of these platforms, 
improving the reproducibility and cross-validation of numerical results, with the ultimate goal of achieving standardization for quantum impurity solvers along the guiding principles of initiatives such as FAIRmat (see \href{https://www.fairmat-nfdi.eu/fairmat/}{fairmat-nfdi.eu/fairmat}).

%\paragraph{closure}
In this paper, we present the mathematical framework, details of the implementation, and key features of \NAME, along with several illustrative applications. We aim to demonstrate how this flexible,
high-performance solver can be used to tackle a wide range of
challenging problems. 
The rest of this work is organized as follows. In \secu{SecInstall} we 
briefly illustrate the overall structure of \NAME together with its dependencies, configuration, and installation procedures. In \secu{SecEDIpack} we present the quantum impurity
problem and review the software implementation in detail. In \secu{SecInterop} we thoroughly discuss the C-binding interface that is at the heart of the 
interoperability capabilities of \NAME. In this section we also discuss 4 interface layers: i) 
the Python API, ii) the TRIQS interface, iii) the W2Dynamics interface and iv) an experimental, yet operative, Julia API.
Finally, in \secu{SecExamples} we discuss the functionalities and the capabilities of \NAME and  all the discussed extensions with a series of illustrative examples.

### 2025-06-02 05_examples.tex
v1.1 arXiv ready AA, MC and GB

In our examples based on the Fortran APIs, we rely on external open-source libraries, such as DMFTtools\footnote{\url{https://github.com/aamaricci/DMFTtools}} (also developed by some of the authors), to perform specific tasks, including evaluating the local Green’s function, implementing the self-consistency, constructing the tight-binding model, or calculating the kinetic energy.
Where appropriate, we include comments that reference these external procedures.

@@

\subsection{Hubbard model on the Bethe lattice (Fortran/C++ API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The description of the Mott transition using the Hubbard model on the Bethe lattice is conventionally regarded as the standard test bed for any DMFT implementation.  \cite{Georges1996RMP,Rozenberg1999PRL,Kotliar1999EPJB,Kotliar2000PRL,Kotliar2002PRL}.

Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw},
$$
where $c^\dagger_{i\s}$ ($c_{i\s}$) are the creation (annihilation) 
operators for an electron at site $i$ with spin $\s$, and 
$n_{i\s} = c^\dagger_{i\s} c_{i\s}$ is the corresponding occupation 
operator. The first sum runs over nearest-neighbor pairs
$\langle ij \rangle$.
We consider the system defined on a Bethe lattice with density of states
$\rho(\e)=\frac{2}{\pi D^2}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
Within DMFT framework \cite{Georges1996RMP}, this lattice model is mapped onto a quantum impurity problem with an effective electronic bath that must be determined self-consistently.

Below, we present the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs. Fully operational versions of the two codes can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, creation of the Bethe DOS and solver initialization.

\begin{center}
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer               :: Le=1000
   real(8)               :: wmixing=0.5d0
   real(8)               :: D=1d0
   integer               :: Nb
   real(8),allocatable   :: Bath(:)
   complex(8),allocatable:: Hloc(:,:,:,:)
   real(8),allocatable   :: Ebands(:),Dbands(:)
   complex(8),allocatable:: Smats(:,:,:,:,:)
   complex(8),allocatable:: Delta(:,:,:,:,:)
   !...  
   !EDIpack: Read variables
   call ed_read_input('inputED.conf')

>>>>

In our examples based on the Fortran APIs, we rely on external open-source libraries, such as DMFTtools (see \href{https://github.com/aamaricci/DMFTtools}{github.com/aamaricci/DMFTtools}), to perform specific tasks, including evaluating the local Green’s function, implementing the self-consistency, constructing the tight-binding model, or calculating the kinetic energy.
Where appropriate, we include comments that reference these external procedures.

@@

\subsection{Hubbard model on the Bethe lattice (Fortran/C++ API, {\tt ed\_mode=normal})}\label{SecExamplesBetheDMFT}
The description of the Mott transition using the Hubbard model on the Bethe lattice is conventionally regarded as the standard test bed for any DMFT implementation.  \cite{Georges1996RMP,Rozenberg1999PRL,Kotliar1999EPJB,Kotliar2000PRL,Kotliar2002PRL}.
%
Here, we present a guided implementation of the DMFT 
solution for the Bethe lattice at $T=0$ using \NAME as impurity
solver.
The model under consideration is the Fermi-Hubbard Hamiltonian:
$$
H = -t \sum_{\langle ij\rangle,\s} c^\dagger_{i\s} c_{j\s} + 
    U \sum_i n_{i\up}n_{i\dw},
$$
where $c^\dagger_{i\s}$ ($c_{i\s}$) are the creation (annihilation) 
operators for an electron at site $i$ with spin $\s$, and 
$n_{i\s} = c^\dagger_{i\s} c_{i\s}$ is the corresponding occupation 
operator. The first sum runs over nearest-neighbor pairs
$\langle ij \rangle$.
We consider the system defined on a Bethe lattice with density of states
$\rho(\e)=\frac{2}{\pi D^2}\sqrt{D^2-\e^2}$,
where $D=2t$ is the half-bandwidth.
Within DMFT framework \cite{Georges1996RMP}, this lattice model is mapped onto a quantum impurity problem with an effective electronic bath that must be determined self-consistently.

Below, we discuss the key components of a basic DMFT implementation 
using \NAME for the Bethe lattice, employing either the Fortran or 
C++ APIs and which can be found 
in the {\tt examples} directory of the \NAME source code. Both 
samples share a similar initial structure, including memory 
allocation, creation of the Bethe DOS and solver initialization:
 \begin{center}
\begin{minipage}[t]{0.49\linewidth}
\textbf{Fortran}
\begin{lstlisting}[style=fstyle,frame=none,numbers=none,basicstyle={\scriptsize\ttfamily}]
program ed_hm_bethe
   USE EDIPACK
   USE SCIFOR
   implicit none
   integer               :: Le=1000
   real(8)               :: wmixing=0.5d0
   real(8)               :: D=1d0
   integer               :: Nb
   real(8),allocatable   :: Bath(:)
   complex(8),allocatable:: Hloc(:,:,:,:)
   real(8),allocatable   :: Ebands(:),Dbands(:)
   complex(8),allocatable:: Smats(:,:,:,:,:)
   complex(8),allocatable:: Delta(:,:,:,:,:)
   !...  
   !EDIpack: Read variables
   call ed_read_input('inputED.conf')

### 2025-06-02 06_conclusions_acknowledgement_appendix.tex
v1.1 arXiv ready AA, MC and GB

\section{Conclusions}
We have presented \NAME{}, a versatile, high-performance solver for
generic quantum impurity problems. Leveraging on the massively
parallel algorithms introduced by its predecessor \cite{Amaricci2022CPC}, this version of
the library features new capabilities and symmetries within a unified
framework, providing for instance reliable evaluations of the dynamical correlation
functions of an arbitrary complex frequency.
This makes it particularly well-suited for capturing the
low-energy excitations and complex spectral features of strongly
correlated systems.
We have shown that this new version enables the evaluation of impurity
reduced density matrices directly from Fock space quantities and thus unlocks 
quantum information analyses of correlated systems, an increasingly
essential tool to address emergent phases of quantum materials.

@@

We thoroughly discussed the implementation of the \NAME library, the
most important algorithms and classes. We presented in
detail the third-party interfaces which extend ED
capabilities beyond the realm of the library itself.   
Finally, we showcased the use of the \NAME software in different
contexts via elaborated examples, which can serve as a reference for potential users.

The modular and extensible design of \NAME provides a
natural foundation for future extensions.
For instance to
cluster-DMFT \cite{Capone2004PRB,Kotliar2006RMP,Park2008PRL} where the
impurity problem is generalized to include clusters of interacting
sites. This approach would enable the treatment of spatial
correlations beyond the single-site approximation, capturing effects
such as d-wave pairing in high-temperature superconductors, charge
order, and complex spin textures in strongly correlated systems.
A different outlook is the development of an integration layer with other
quantum embedding methods, such ghost Gutzwiller \cite{Lanata2015PRX},
which heavily rely on solving effective impurity problems.    
We anticipate that \NAME will become a valuable tool for the
computational condensed matter community, supporting a wide range of
studies on strongly correlated materials and providing a reliable
reference platform for quantum impurity solvers.

@@

\paragraph{Author contributions}
A.A. designed the project, developed the main structure of the software and conceived the computational algorithms in the core library. 
L.C., G.M., F.P. A.S., S.G. and G.B.  equally contributed to the development and implementation of many different aspects of the  software, including code writing, configuration, installation, testing and continuous integration procedures. 
L.C. devised and developed the C-binding layer, the Python, Julia and Monicelli interface together with Python specific deployment and installation tools. 
I.K. developed the TRIQS interface and contributed to the testing part. 
A.K., M.W. and G.S. developed the integration with W2Dynamics.  
A.A. and M.C. conceived the diagonalization, correlation function  calculation and bath optimization strategies implemented in the library. M.C. devised the inclusion of the electron-phonon coupling. 
A.A., M.C., L.d.M. and G.S designed additional algorithms for extended functionalities of the core software.   
A.A., L.C., I.K., A.K., S.G. and G.B. performed the calculations, analyzed the results and prepared the data presentation in the examples.
A.A. and L.C. wrote the backbone of the manuscript.
All authors participated in writing this manuscript.

>>>>

\section{Conclusions}
We have presented \NAME{}, a versatile and high-performance ED solver for generic quantum impurity problems. Building on the massively
parallel algorithms introduced by its predecessor \cite{Amaricci2022CPC}, this version of
the library features new capabilities allowing for different broken-symmetry solutions within a unified
framework that provides, for instance, reliable evaluations of a variety of local dynamical correlation
functions of an arbitrary complex frequency ranging from the Green's functions to superconducting, excitonic and magnetic response functions.

This new version allows for an efficient evaluation of 
reduced density matrices of the impurity directly from Fock space quantities and thus enables  
quantum information analysis of correlated systems, which is becoming an essential tool to address emergent phases of quantum materials.

@@

We thoroughly discussed the implementation of the \NAME library and its
most important algorithms and classes. We presented in
detail the third-party interfaces which extend ED
capabilities beyond the domain of the library itself. 
Finally, we showcased the use of the \NAME software in different
contexts through elaborated examples, which can serve as a reference for potential users.

The modular and extensible design of \NAME provides a
natural foundation for future extensions.
For instance it provides a suitable basis to implement cluster-DMFT \cite{Capone2004PRB,Kotliar2006RMP,Park2008PRL} methods, where the
impurity problem is generalized to include clusters of interacting
sites. This approach includes spatial correlations beyond the single-site approximation, capturing effects
such as $d$-wave pairing, charge ordering, and complex spin textures in strongly correlated systems.
A different outlook is the development of an integration layer with other quantum embedding methods, such as the ghost Gutzwiller scheme \cite{Lanata2015PRX},
which heavily rely on solving effective impurity problems.    
We anticipate that \NAME will become a valuable tool for the computational condensed matter community, supporting a wide range of
studies on strongly correlated materials and providing a reliable
reference platform for quantum impurity solvers.

@@

\paragraph{Author contributions}
A.A. designed the project, developed the main structure of the software and conceived the computational algorithms in the core library. A.A. and M.C. conceived the diagonalization, correlation function  calculation and bath optimization strategies implemented in the library. M.C. devised the inclusion of the electron-phonon coupling and of the superconducting state. 
L.C., G.M., F.P. A.S., S.G. and G.B.  equally contributed to the development and implementation of many different aspects of the  software, including code writing, configuration, installation, testing and continuous integration procedures. 
L.C. devised and developed the C-binding layer, the Python, Julia and Monicelli interface together with Python specific deployment and installation tools. 
I.K. developed the TRIQS interface and contributed to the testing part. 
A.K., M.W. and G.S. developed the integration with W2Dynamics.  
A.A., M.C., L.d.M. and G.S designed additional algorithms for extended functionalities of the core software.   
A.A., L.C., I.K., A.K., S.G. and G.B. performed the calculations, analyzed the results and prepared the data presentation in the examples.
A.A. and L.C. wrote the backbone of the manuscript.
All authors participated in writing this manuscript.
